---
title: "Predicting Housing Code Violations"
author: "Maxwell Austensen"
date: '`r Sys.Date()`'
output:
  html_notebook: default
  github_document: default
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
# If you cloned this repo but haven't run "make_data.R" then you'll only have 
# A zipped version of this file, so it needs to be unzipped first
if (!exists("../data/merged.feather")) {
  unzip("../data/merged.zip", files = "data/merged.feather", exdir = "../data")
  file.rename("../data/data/merged.feather", "../data/merged.feather")
  file.remove("../data/data")
}

df <- feather::read_feather("../data/merged.feather")

```

Small properties represent a very large share of all properties and contain a large share of all residential units.

```{r}
df %>% 
  group_by(res_units) %>% 
  summarise(sh_all_bbls = n() / nrow(.)) %>%  
  filter(res_units <= 50) %>% 
  ggplot(aes(res_units, sh_all_bbls)) + 
  geom_col() + 
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Share of All Properties with 3+ Units in 2016, by Property Size",
       subtitle = "Graph truncated at 50 residential units",
       y = NULL, x = "Number of Residential Units in Property")
```


```{r}
all_res_units <- sum(df$res_units)

df %>% 
  group_by(res_units) %>% 
  summarise(sh_all_units = sum(res_units) / all_res_units) %>%  
  filter(res_units <= 200) %>%
  ggplot(aes(res_units, sh_all_units)) + 
  geom_col(width = 1) + 
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Share of All Residential Units in 3+ Unit Properties in 2016, by Property Size",
       subtitle = "Graph truncated at 200 residential units",
       y = NULL, x = "Number of Residential Units in Property")
```

```{r}
df %>% 
  filter(res_units <= 200) %>% 
  group_by(res_units) %>%
  summarise(any_serious_viol = mean(viol_bbl_ser_2016 > 0)) %>% 
  filter(!is.na(res_units)) %>% 
  ggplot(aes(res_units, any_serious_viol)) + 
  geom_point() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Share of Properties with Any Serious Violations in 2016, by Property Size",
       subtitle = "Graph truncated at 200 residential units",
       y = NULL, x = "Number of Residential Units in Property")
```

To get an overall score for each building we might want to adjust the violation counts for building size. Violations that affect a specific apartment are adjusted by the number of residential units, and violations that affect the entire building are not adjusted.

```{r}
df %>% 
  filter(res_units <= 200) %>% 
  group_by(res_units) %>%
  summarise(avg_viol_ser_all_2016 = mean(viol_bbl_ser_2016, na.rm = TRUE)) %>% 
  ggplot(aes(res_units, avg_viol_ser_all_2016)) + 
  geom_point() +
  labs(title = "Average Number of Adjusted Serious Violations in 2016, by Property Size",
       subtitle = "Graph truncated at 200 residential units",
       y = NULL, x = "Number of Residential Units in Property")
```

```{r}
df %>% 
  filter(viol_bbl_ser_2016 > 0) %>% 
  ggplot(aes(viol_bbl_ser_2016)) + 
  geom_density() +
  labs(title = "Density of Adjusted Number of Serious Violations in 2016",
       subtitle = "Among properties with at least one serious violation in 2016",
       y = NULL, x = "Adjusted Number of Serious Violations")
```

```{r}
sh_viol_16 <- mean(df$viol_bbl_ser_2016 > 0)

viol_16_avg <- df %>% 
  filter(viol_bbl_ser_2016 > 0) %>% 
  summarise(mean(viol_bbl_ser_2016)) %>%
  .[[1]]
 
viol_16_no15 <- df %>% 
  filter(viol_bbl_ser_2016 > 0) %>% 
  summarise(mean(viol_bbl_ser_2015 > 0)) %>%
  .[[1]]
```

Only `r round(sh_viol_16*100, 1)`% of all properties with 3+ units had any serious violations in 2016. Among these properties that had any serious violations, the average number of adjusted serious violations was `r round(viol_16_avg, 1)` in 2016. Of all the those properties that had at least one serious violation in 2016, `r round(viol_16_no15*100, 1)`% didn't have a serious violation in 2015.




